Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-02-08 constantRate/ConstCompFuzzExt.tex
minor cleanup

\defref{def:partial source} is a weakening of block sources~(introduced by Chor and Goldreich~\cite{DBLP:journals/siamcomp/ChorG88}) as only some blocks are required to have entropy conditioned on the past.  The need for remaining entropy conditioned on the past is arbitrary.  The actual requirement is that there exists some ordering of indices where most items have entropy conditioned on all previous items in this ordering~(for example, a ``partial'' reverse block source~\cite{vadhan2003constructing}).
%Note a partial block source requires worst case entropy while an unguessable block distribution only requires average conditional entropy. \lnote{this comparison is unclear to me}  However, we achieve security for significantly lower entropy levels than for unguessable block distributions.

We use an algorithm that selects a fixed size random subset of $\{1,..., \gamma\}$.  Let $\sample_{\gamma, \eta}(\cdot)$ be an algorithm that outputs a random subset of $\{1,..., \gamma\}$ of size $\eta$ and let $r_{sam}$ the number of required coins for $\sample_{\gamma, \eta}$.

@@

The main change in \consref{cons:sampling} is that the obfuscated values are concatenated symbols of $W$.  This paradigm is similar to \emph{sample-then-extract} from the locally computable extractors literature~\cite{lu2002hyper,vadhan2003constructing}.  For this reason we call \consref{cons:sampling} \emph{sample-then-obfuscate}.  A crucial difference is that the use of a computational primitive~(obfuscation) allows us to sample multiple times, because we need to argue only about individual entropy of $v_i$, as opposed to the information-theoretic setting, where it would be necessary to argue about the joint entropy of the $v_i$ values.

@@

\subsection{Discussion of \consref{cons:sampling}}
We now show \consref{cons:sampling} can work for partial block sources when $\Huse\le 0$.  The required entropy of partial block source is $\alpha (\gamma-\beta ) = \Theta(\gamma)$.  We are able to correct $O(\gamma/\eta)$ errors.
This yields:
\begin{align*}
\Huse &= \Hoo(W) -\log |B_t| \\
&< \Theta(\gamma)- t \log |\mathcal{Z}|\\
&= \Theta(\gamma) - \Theta(\gamma/\eta) \log |\mathcal{Z}|
\end{align*}
That is, \consref{cons:sampling} achieves $\Huse\le 0$ when the starting alphabet is super polynomial~(noting that for super polynomial size $\mathcal{Z}$ we can set $\eta$ to be super logarithmic and $o(\log |\mathcal{Z}|)$).  We note that for polynomial-size alphabets, \consref{cons:sampling} will still work as long as we use a code that corrects Hamming errors in both directions (with a polynomial size alphabet the probability of $1\rightarrow 0$ bits flips is noticeable); however, for a polynomial-size alphabet, $\Huse>0$.
%However, in this setting $\Huse =\Omega(\gamma)$ and known information-theoretic fuzzy extractors provide superior performance.
\bibliographystyle{alpha}
\bibliography{crypto}

@@

\section{Fuzzy Conductors}

\label{sec:conductors}
Fuzzy extractors  have strong upper bounds on remaining entropy based on the best error correcting codes~(if they provide the same guarantee for all input distributions with the same entropy and error tolerance).  Fuller, Meng, and Reyzin show that computational information-reconciliation techniques are subject to similar bounds~\cite{fuller2013computational}.  They suggest these bounds may be avoided by outputting a fresh random variable.  This is known as a fuzzy conductor~\cite{KanukurthiR09}.
\begin{definition}
A $(\mathcal{M}, m, \tilde{m}, t, \epsilon)$-\emph{fuzzy conductor} with error $\delta$ is a pair of randomized procedures, ``generate''~(\gen') and ``reproduce''~(\rep'), with the following properties:
\begin{enumerate}
\item The generate procedure $\gen'$ on input $w\in \mathcal{M}$ outputs a string $x\in\zo^*$ and a helper string $ss\in\zo^*$.
\item The reproduction procedure $\rep'$ takes an element $w'\in\mathcal{M}$ and a bit string $ss\in\zo^*$ as inputs.  The \emph{correctness} property of fuzzy extractors guarantees that for $w$ and $w'$ such that $\dis(w, w')\leq t$, if $X, SS$ were generated by $(X, SS)\leftarrow \gen'(w)$, then $\Pr[\rep(w', SS) = X]\geq 1-\delta$~(over the coins of $\gen', \rep'$).  If $\dis(w, w')>t$, then no guarantee is provided about the output of $\rep'$.
\item The security property guarantees that for any distribution $W$ on $\mathcal{M}$ of min-entropy $m$, the string $X$ is close to a high entropy distribution $Y$.  That is, $\exists Y$ with $\Hav(Y | SS ) \geq \tilde{m}$ such that $\Delta((X, SS), (Y, SS))\leq \epsilon$.
\end{enumerate}
\end{definition}

@@

\section{Proof of \lemref{lem:cond and cext}}
\label{sec:cond and cext}
\begin{proof}
It suffices to show if there is some distinguisher $D'$ of size $s'$ where
\[\delta^{D'}((\cext(X; U_d), U_d, SS), (U_\kappa, U_d, SS))>\epsilon_{cond}+ \epsilon_{ext}\]
 then there is an distinguisher $D$ of size $s_{cond}$ where for all $Y$ where $\Hav(Y|SS) \geq \tilde{m}$ such that
 \[
 \delta^{D}((X, SS), (Y, SS))\geq \epsilon_{cond}.
 \]
Let $D'$ be such a distinguisher.  That is,
\[
\delta^{D'}(\cext(X, U_d)\times U_d \times SS, U_\kappa\times U_d\times SS)> \epsilon_{ext}+\epsilon_{cond}.
\]
Then define $D$ as follows.  On input $(y, ss)$ sample $seed\leftarrow U_d$, compute $r\leftarrow \cext(y; seed)$ and output $D(r, seed, ss)$.  Note that $|D| \approx s' + |\cext| +d= s_{cond}$.  Then we have the following:
\begin{align*}
\delta^{D}((X, SS), (Y, SS))&= \delta^{D'}((\cext(X, U_d), U_d, SS), \cext(Y, U_d), U_d, SS)\\
&\geq \delta^{D'}((\cext(X, U_d), U_d, SS), (U_\kappa\times U_d \times SS)) - \delta^{D'}((\cext(Y, U_d), U_d, SS), (U_\kappa\times U_d \times SS))\\
&>\epsilon_{cond}+\epsilon_{ext}- \epsilon_{ext} = \epsilon_{cond}.
\end{align*}
Where the last line follows by noting that $D'$ is of size at most $s_{ext}$.  Thus $D$ distinguishes $X$ from all $Y$ with sufficient conditional min-entropy.  This is a contradiction.
\end{proof}

\section{Proof of \lemref{lem:security of cons}}
\label{app:security of main cons}
\begin{proof}

Let $\mathcal{O}$ be a $\ell$-composable VGB obfuscator with auxiliary input for point programs over $\mathcal{Z}$.  Let $W$ be a $(q, \alpha = \omega(\log n), \beta)$-unguessable block distribution.  Our goal is to show that for all $s_{sec} = \poly(n)$ there exists $\epsilon_{sec} =\ngl(n)$ such that $H^{\hill}_{\epsilon_{sec}, s_{sec}}(C|P)\geq H_0(C)- \beta$. % for $\epsilon' = 2\epsilon_{obf} + (\ell - \beta)2^{-(\alpha-1)}$.
Suppose not, that is suppose there is some $s_{sec} = \poly(n)$ such that exists $\epsilon_{sec} = \poly(n)$ and $H^{\hill}_{\epsilon_{sec}, s_{sec}}(C|P) < H_0(C)-\beta$.
\lnote{some $i$s should be changed to $j$s below this line, but I ran out of time; will do}
By \defref{def:block guessable} there exists a set of indices $J$ such that all blocks within $J$ are unguessable.  Define by $C'$ the distribution of sampling a uniform codeword where all locations outside $J$ are fixed.  Then $C | C_{J^c} \overset{d}=C'$.  By \clref{cl:many locations ent}, we have that $\Hav(C|C_{J^c} )= H_0(C) -\beta$ and thus $\Hav(C'| C_{J^c}) = H_0(C) -\beta$.
Let $D$ a distinguisher of size be some distinguisher of size at most $s_{sec}$ such that
\[
| \expe[D(C, P)] - \expe[D(C', P)] > \epsilon_{sec} = 1/\poly(n).
\]
Define the distribution $X$ as follows:
\[X_j =
\begin{cases}
W_j & C_j = 0\\
R_j & C_j = 1.
\end{cases}\]  By the security of obfuscation~(\defref{def:obf}), there exists a unbounded time simulator $S$~(making at most $q$ queries) such that
\begin{align}
\label{eq:dist before}
|\expe [D(P_1,..., P_\ell, C)] - \expe [S^{I_X(\cdot, \cdot)}(C, 1^{\ell \log |Z|})] |\leq \epsilon_{sec}/3.
\end{align}
We now prove $S$ cannot distinguish between $C$ and $C'$.
\begin{lemma}
\label{lem:sim cannot distinguish}
$\Delta(S^{I_X(\cdot, \cdot)}(C, 1^{\ell \log |Z|}), S^{I_X(\cdot, \cdot)}(C', 1^{\ell \log |Z|})) \le (\ell-\beta) 2^{-(\alpha+1)}$.
\end{lemma}

>>>>

\defref{def:partial source} is a weakening of block sources~(introduced by Chor and Goldreich~\cite{DBLP:journals/siamcomp/ChorG88}), as only some blocks are required to have entropy conditioned on the past.  The choice of conditioning on the past is arbitrary: a more general sufficient condition is that there exists some ordering of indices where most items have entropy conditioned on all previous items in this ordering~(for example, a ``partial'' reverse block source~\cite{vadhan2003constructing}).
%Note a partial block source requires worst case entropy while an unguessable block distribution only requires average conditional entropy. \lnote{this comparison is unclear to me}  However, we achieve security for significantly lower entropy levels than for unguessable block distributions.

Let $\sample_{\gamma, \eta}(\cdot)$ be an algorithm that  outputs a random subset of $\{1,..., \gamma\}$ of size $\eta$ given let $r_{sam}$ bits of randomness.

@@

The main change in \consref{cons:sampling} is that the obfuscated values are concatenated symbols of $W$.  This paradigm is similar to \emph{sample-then-extract} from the locally computable extractors literature~\cite{lu2002hyper,vadhan2003constructing}.  For this reason we call \consref{cons:sampling} \emph{sample-then-obfuscate}.  A crucial difference is that the use of a computational primitive~(obfuscation) allows us to sample multiple times, because we need to argue only about individual entropy of $V_i$, as opposed to the information-theoretic setting, where it would be necessary to argue about the entropy of the joint variable $V$.

@@

\subsection{Discussion of \consref{cons:sampling}}
\lnote{I need to check this}
We now show \consref{cons:sampling} can work for partial block sources when $\Huse\le 0$.  The required entropy of partial block source is $\alpha (\gamma-\beta ) = \Theta(\gamma)$.  We are able to correct $O(\gamma/\eta)$ errors.
This yields:
\begin{align*}
\Huse &= \Hoo(W) -\log |B_t| \\
&< \Theta(\gamma)- t \log |\mathcal{Z}|\\
&= \Theta(\gamma) - \Theta(\gamma/\eta) \log |\mathcal{Z}|
\end{align*}
That is, \consref{cons:sampling} achieves $\Huse\le 0$ when the starting alphabet is super polynomial~(noting that for super polynomial size $\mathcal{Z}$ we can set $\eta$ to be super logarithmic and $o(\log |\mathcal{Z}|)$).  We note that for polynomial-size alphabets, \consref{cons:sampling} will still work as long as we use a code that corrects Hamming errors in both directions (with a polynomial size alphabet the probability of $1\rightarrow 0$ bits flips is noticeable); however, for a polynomial-size alphabet, $\Huse>0$.
%However, in this setting $\Huse =\Omega(\gamma)$ and known information-theoretic fuzzy extractors provide superior performance.
\bibliographystyle{alpha}
\bibliography{crypto}

@@

\section{Fuzzy Conductors}
\lnote{I need to check this}
\label{sec:conductors}
Fuzzy extractors  have strong upper bounds on remaining entropy based on the best error correcting codes~(if they provide the same guarantee for all input distributions with the same entropy and error tolerance).  Fuller, Meng, and Reyzin show that computational information-reconciliation techniques are subject to similar bounds~\cite{fuller2013computational}.  They suggest these bounds may be avoided by outputting a fresh random variable.  This is known as a fuzzy conductor~\cite{KanukurthiR09}.
\begin{definition}
A $(\mathcal{M}, m, \tilde{m}, t, \epsilon)$-\emph{fuzzy conductor} with error $\delta$ is a pair of randomized procedures, ``generate''~(\gen') and ``reproduce''~(\rep'), with the following properties:
\begin{enumerate}
\item The generate procedure $\gen'$ on input $w\in \mathcal{M}$ outputs a string $x\in\zo^*$ and a helper string $ss\in\zo^*$.
\item The reproduction procedure $\rep'$ takes an element $w'\in\mathcal{M}$ and a bit string $ss\in\zo^*$ as inputs.  The \emph{correctness} property of fuzzy extractors guarantees that for $w$ and $w'$ such that $\dis(w, w')\leq t$, if $X, SS$ were generated by $(X, SS)\leftarrow \gen'(w)$, then $\Pr[\rep(w', SS) = X]\geq 1-\delta$~(over the coins of $\gen', \rep'$).  If $\dis(w, w')>t$, then no guarantee is provided about the output of $\rep'$.
\item The security property guarantees that for any distribution $W$ on $\mathcal{M}$ of min-entropy $m$, the string $X$ is close to a high entropy distribution $Y$.  That is, $\exists Y$ with $\Hav(Y | SS ) \geq \tilde{m}$ such that $\Delta((X, SS), (Y, SS))\leq \epsilon$.
\end{enumerate}
\end{definition}

@@

\section{Proof of \lemref{lem:cond and cext}}
\label{sec:cond and cext}
\lnote{I need to check this}
\begin{proof}
It suffices to show if there is some distinguisher $D'$ of size $s'$ where
\[\delta^{D'}((\cext(X; U_d), U_d, SS), (U_\kappa, U_d, SS))>\epsilon_{cond}+ \epsilon_{ext}\]
 then there is an distinguisher $D$ of size $s_{cond}$ where for all $Y$ where $\Hav(Y|SS) \geq \tilde{m}$ such that
 \[
 \delta^{D}((X, SS), (Y, SS))\geq \epsilon_{cond}.
 \]
Let $D'$ be such a distinguisher.  That is,
\[
\delta^{D'}(\cext(X, U_d)\times U_d \times SS, U_\kappa\times U_d\times SS)> \epsilon_{ext}+\epsilon_{cond}.
\]
Then define $D$ as follows.  On input $(y, ss)$ sample $seed\leftarrow U_d$, compute $r\leftarrow \cext(y; seed)$ and output $D(r, seed, ss)$.  Note that $|D| \approx s' + |\cext| +d= s_{cond}$.  Then we have the following:
\begin{align*}
\delta^{D}((X, SS), (Y, SS))&= \delta^{D'}((\cext(X, U_d), U_d, SS), \cext(Y, U_d), U_d, SS)\\
&\geq \delta^{D'}((\cext(X, U_d), U_d, SS), (U_\kappa\times U_d \times SS)) - \delta^{D'}((\cext(Y, U_d), U_d, SS), (U_\kappa\times U_d \times SS))\\
&>\epsilon_{cond}+\epsilon_{ext}- \epsilon_{ext} = \epsilon_{cond}.
\end{align*}
Where the last line follows by noting that $D'$ is of size at most $s_{ext}$.  Thus $D$ distinguishes $X$ from all $Y$ with sufficient conditional min-entropy.  This is a contradiction.
\end{proof}

\section{Proof of \lemref{lem:security of cons}}
\label{app:security of main cons}
\begin{proof}
\lnote{I need to check this}
Let $\mathcal{O}$ be a $\ell$-composable VGB obfuscator with auxiliary input for point programs over $\mathcal{Z}$.  Let $W$ be a $(q, \alpha = \omega(\log n), \beta)$-unguessable block distribution.  Our goal is to show that for all $s_{sec} = \poly(n)$ there exists $\epsilon_{sec} =\ngl(n)$ such that $H^{\hill}_{\epsilon_{sec}, s_{sec}}(C|P)\geq H_0(C)- \beta$. % for $\epsilon' = 2\epsilon_{obf} + (\ell - \beta)2^{-(\alpha-1)}$.
Suppose not, that is suppose there is some $s_{sec} = \poly(n)$ such that exists $\epsilon_{sec} = \poly(n)$ and $H^{\hill}_{\epsilon_{sec}, s_{sec}}(C|P) < H_0(C)-\beta$.
By \defref{def:block guessable} there exists a set of indices $J$ such that all blocks within $J$ are unguessable.  Define by $C'$ the distribution of sampling a uniform codeword where all locations outside $J$ are fixed.  Then $C | C_{J^c} \overset{d}=C'$.  By \clref{cl:many locations ent}, we have that $\Hav(C|C_{J^c} )= H_0(C) -\beta$ and thus $\Hav(C'| C_{J^c}) = H_0(C) -\beta$.
Let $D$ a distinguisher of size be some distinguisher of size at most $s_{sec}$ such that
\[
| \expe[D(C, P)] - \expe[D(C', P)] > \epsilon_{sec} = 1/\poly(n).
\]
Define the distribution $X$ as follows:
\[X_j =
\begin{cases}
W_j & C_j = 0\\
R_j & C_j = 1.
\end{cases}\]  By the security of obfuscation~(\defref{def:obf}), there exists a unbounded time simulator $S$~(making at most $q$ queries) such that
\begin{align}
\label{eq:dist before}
|\expe [D(P_1,..., P_\ell, C)] - \expe [S^{I_X(\cdot, \cdot)}(C, 1^{\ell \log |Z|})] |\leq \epsilon_{sec}/3.
\end{align}
We now prove $S$ cannot distinguish between $C$ and $C'$.
\begin{lemma}
\label{lem:sim cannot distinguish}
$\Delta(S^{I_X(\cdot, \cdot)}(C, 1^{\ell \log |Z|}), S^{I_X(\cdot, \cdot)}(C', 1^{\ell \log |Z|})) \le (\ell-\beta) 2^{-(\alpha+1)}$.
\end{lemma}
